---
author: "Carl Boettiger"
institute: "UC Berkeley"
date: "`r Sys.Date()`"
title: "Content-based Identifiers for Iterative Forecasts: A Proposal"
output:
  xaringan::moon_reader:
    css: ["default", "solarized-light.css"]
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: "solarized-light"
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(#dev.args=list(bg="transparent"), 
                      message=FALSE, 
                      warning=FALSE,
                      fig.width=8, 
                      fig.height=5, 
                      cache = TRUE)

library(tidyverse)

library(icon)
library(ggthemes)
library(magick)
library(gganimate)
library(animation)
library(patchwork)



theme_set(theme_solarized(base_size=16))
scale_colour_discrete <- function(...) scale_colour_solarized()
scale_fill_discrete <- function(...) scale_fill_solarized()
pal <- solarized_pal()(6)
txtcolor <- "#586e75"

ggimage <- function(path, transparent = FALSE){
  img <- magick::image_read(path)
  if(transparent)  img <- magick::image_transparent(img, "white")
  ggplot2::ggplot() + 
  ggplot2::annotation_raster(img, 
                    xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
    theme(axis.line = element_blank(), 
          panel.background = element_blank(),
          axis.ticks = element_blank(), 
          axis.text = element_blank(), 
          axis.title = element_blank())
}
ggblank <- function() ggplot() + geom_blank() + 
  theme(axis.line = element_blank(), panel.background = element_blank())

```

layout: true
background-color: #fdf6e3
class: center, top

---

# Content-Based Identifiers For Iterative Forecasts: 

## A Proposal

<div class="my-footer">

<a href="https://carlboettiger.info"> `r icon::fa("user")` Carl Boettiger</a> | 
<a href="https://berkeley.edu"> `r icon::fa("briefcase")` UC Berkeley</a> | 
<a href="https://twitter.com/cboettig"> `r icon::fa("twitter")` @cboettig</a>

</div>


---

# Most scientific scripts have to read in data.

---

# In R, sometimes we read data in like this:

```{r}
df <- 
read_tsv("/home/rstudio/Documents/cboettig/talks-2020-dataone/vostok.icecore.co2")
```

---

# A local path is a bit more portable...

```{r}
df <- read_tsv("vostok.icecore.co2")
```

--

## ... if you have the data file! 

---

# Reading from URLs gets around this...

```{r}
url <- paste0("https://knb.ecoinformatics.org/", 
              "knb/d1/mn/v2/object/", 
              "ess-dive-457358fdc81d3a5-20180726T203952542")
df <- read_tsv(url)
```

--

## ... but we don't want to re-download every time we re-run

---

# Caching a copy can help...

```{r}
library(pins)
pins::pin(url)
```

--

## ... but links still rot

---

# DOIs can solve link rot


## but you can almost never put DOIs into scripts...

```{r}
download.file("https://doi.org/10.3334/CDIAC/ATG.009",
              "vostok.icecore.co2")
```


---

# So what should we do?

---


# Two observations:

#### Reproducible code must read in the same bytes every time
#### Checksums, such as `SHA-256`, are the gold standard way to confirm that.

---

# A proposal

## What if we could read in content by its SHA-256 hash?

```{r}
library(contentid)
content_id("vostok.icecore.co2")
```

---

```{r}
file <- resolve("hash://sha256/9412325831dab22aeebdd674b6eb53ba6b7bdd04bb99a4dbb21ddff646287e37")
content_id(file)
```

---

## whoops, long hashes are liable to be cut off

```{r}
file <- resolve("hash://sha256/9412325831dab22aeebdd674b")
```
## Fortunately, that's not a problem!

```{r}
content_id(file)
```

Recall that hashes encode the most significant bytes first

---


# So where is the data coming from?

```{r}
sources <- contentid::query_sources("hash://sha256/9412325831dab22aeebdd674b6eb53ba6b7bdd04bb99a4dbb21ddff646287e37")
sources
```

---

# Content-based identifers are 
# location agnostic

---

# LOCKSS:

## Lots of copies keeps stuff safe

---

# How?

---

# Registering a location:

```{r}
url <- paste0("https://knb.ecoinformatics.org/", 
              "knb/d1/mn/v2/object/", 
              "ess-dive-457358fdc81d3a5-20180726T203952542")
register(url)
```

---

# Existing Registeries

## DataONE 

If you publish to DataONE using SHA-256 sum, or know data uses it, you can query for it.

## GitHub / Software Heritage

Trigger a Software Heritage API of a repository, query content by hash


---

# But WAIT! 

# I only want the latest data!

---

# Access the latest data

(Mauna Loa CO2 data, updated monthly)


```{r}
download.file("ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt",
              "co2_mm_mlo.txt")
content_id("co2_mm_mlo.txt")
```

---

Lightweight provenance tracing


```{r}
# remotes::install_github("cboettig/prov")
library(prov)
write_prov("co2_mm_mlo.txt")
```


----

Tracking input and output data

```{r}
read_table("co2_mm_mlo.txt", comment = "#", 
         col_names = c("", "", "date", "co2", "", "", "", "")) %>%
  ggplot(aes(date, co2)) + geom_line()
ggsave("co2.png")
```

---

```{r}
write_prov(data_in = "co2_mm_mlo.txt", 
           data_out = "co2.png",
           code = "index.Rmd")
```

---

# Richer examples: an iterative forecast

- [Identifiers for Iterative Forecasts](https://cboettig.github.io/contentid/articles/technotes/identifiers-for-iterative-forecasting.html)
- <https://github.com/eco4cast/neon4cast-beetles>
- [JSON-LD Playground](https://tinyurl.com/y58no3fq)
